---
output: github_document
---

```{r echo=FALSE, results = 'asis'}
pkg <- 'streamConnect'

source("https://raw.githubusercontent.com/mhahsler/pkg_helpers/main/pkg_helpers.R")
pkg_title(pkg)
```

The R package is part of the  [stream](https://github.com/mhahsler/stream) ecosystem. It 
adds functionality to connect stream mining components from package stream using sockets and Web services. The package can be used create distributed workflows and create plumber-based Web services which can be deployed on most common cloud services.

```{r echo=FALSE, results = 'asis'}
pkg_citation(pkg)
```

```{r echo=FALSE, results = 'asis'}
pkg_install(pkg)
```

## Examples

See [stream: Working With Data Streams using Connections and Web Services](https://CRAN.R-project.org/package=streamConnect/vignettes/connections.html)
for examples.

## Acknowledgements

The development of the stream package was supported in part by NSF CMMI 1728612.

## References

Michael Hahsler, Matthew Bolaños, and John Forrest. [stream: An extensible framework for data stream clustering research with R.](https://dx.doi.org/10.18637/jss.v076.i14) _Journal of Statistical Software,_ 76(14), February 2017.
