---
output: github_document
---

```{r echo=FALSE, results = 'asis'}
pkg <- 'streamConnect'

source("https://raw.githubusercontent.com/mhahsler/pkg_helpers/main/pkg_helpers.R")
pkg_title(pkg)
```

The R package is part of the [stream](https://michael.hahsler.net/stream) ecosystem. It
adds functionality to connect stream mining components from the `stream` package using sockets and Web services. The package can be used to create distributed workflows and plumber-based Web services that can be deployed on most common cloud services.

```{r echo=FALSE, results = 'asis'}
pkg_citation(pkg)
```

```{r echo=FALSE, results = 'asis'}
pkg_install(pkg)
```

## Examples

See the [Getting Started with streamConnect](https://michael.hahsler.net/streamConnect/articles/streamConnect.html)
vignette for examples.

## Acknowledgements

The development of the stream package was supported in part by NSF CMMI 1728612.

## References

Michael Hahsler, Matthew Bolaños, and John Forrest. [stream: An extensible framework for data stream clustering research with R.](https://dx.doi.org/10.18637/jss.v076.i14) _Journal of Statistical Software,_ 76(14), February 2017.
