---
title: "stream: Working With Data Streams using Connections and Web Services"
author: "Michael Hahsler"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{stream: Working With Data Streams using Connections and Web Services}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Data streams are often processed in a distributed manner
using multiple machines or multiple processes.
For example, a data stream may be produced by a sensor attached to a remote 
machine or multiple clustering algorithms run in parallel using several R 
processes. Another application is to connect to other software components 
in a stream mining pipeline.

First, we show how socket connections together with the package `stream` can be used to 
connect multiple processes or machines. 

Then we give examples of how
package `streamConnect` makes connecting stream mining components more
convenient by providing an interface
to connect stream processing using _sockets_ or _web services_. 
While sockets are only used to connect data steam generating processes,
web services are more versatile and can also be used
to create data stream clustering processes as a service.

The last section of this paper shows how to deploy the server/web service.

For the examples below, we will use a random available port.

```{r}
port <- httpuv::randomPort()
port
```

# Using Sockets Directly to Publish a Data Stream

The functions `write_stream()` and the 
class `DSD_ReadStream` provided in package `stream` can be used for communicate 
via connections (files, sockets, URLs, etc.). 
In the first example, we manually set up the connection. The example
is useful to understand how sockets work especially for users interested in 
implementing their own components using other programming languages or 
connecting with other data stream software.

A more convenient way
to do this using package `streamConnect` is described later in this paper.

## Server: Serving a Data Stream

The server serves data from a data stream.
We use library `callr` to create a separate R process that serves a data 
stream creating 10 points every second
using a socket connection, but
you can also put the code in function `r_bg()` in a file 
called `server.R` and run (potentially on a different machine) 
it with `R CMD BATCH server.R` from the command line.

```{r eval = TRUE}
library(stream)
library(callr)

rp1 <- r_bg(function(port) {
  library(stream)
  stream <- DSD_Gaussians(k = 3, d = 3)
  blocksize <- 10

  con <- socketConnection(port = port, server = TRUE)
  
  while (TRUE) {
    write_stream(stream, con, n = blocksize, close = FALSE)
    Sys.sleep(1)
  }
  
  close(con)
}, 
  args = list(port = port))

rp1
```

## Client: Reading from the Stream

The client consumes the data stream.
We open the connection which starts the data generating process. 
Note that `streamConnect` is not used here. For convenience, we only use the helper
`retry()` defined in streamConnect to make sure the server connections are established.


```{r eval = TRUE}
con <- streamConnect::retry(socketConnection(port = port, open = 'r'))
con

dsd <- streamConnect::retry(DSD_ReadStream(con))
```

We poll
all available data (`n = -1`) several times. The first request should yield 
10 points, the second none and the third request should 
yield 20 points (2 seconds).

```{r eval = TRUE}
get_points(dsd, n= -1)
get_points(dsd, n= -1)

Sys.sleep(2)
get_points(dsd, n= -1)

close(con)
```

## Server: Stoping the Server Process

Here we stop the `callr` process. Note that the socket connection is still active and will serve the data in the connection buffer as long as the reading
process keeps the connection open.

```{r eval = TRUE}
rp1$kill()
```

# streamConnect Sockets

`streamConnect` provides a more convenient way to set up a connection using sockets. 
`publish_DSD_via_Socket()` creates a socket broadcasting the data stream and
`DSD_ReadSocket` creates a `DSD` object reading from that socket.

## Server: Publish Data
We create a DSD process sending data to the port.
```{r eval = TRUE}
library(streamConnect)

rp1 <- DSD_Gaussians(k = 3, d = 3) %>% publish_DSD_via_Socket(port = port)
rp1
```

## Client: Connect to the Data Stream
Next, we create a DSD that connects to the socket. `DSD_ReadSocket()` already performs internally 
retries 
```{r eval = TRUE}
library(streamConnect)

dsd <- DSD_ReadSocket(port = port, col.names = c("x", "y", "z", ".class"))
dsd

get_points(dsd, n = 10)
plot(dsd)

close_stream(dsd)
```

## Server: Stoping the Server Process
Closing the stream on the client also closes the connection which may already 
kill the serving process.
```{r eval = TRUE}
if (rp1$is_alive()) rp1$kill()
```


# streamConnect Web Services

Web services are more versatile, they can be used to deploy data stream sources 
using `publish_DSD_via_WebService()`/`DSD_ReadWebservice`
or data stream tasks using `publish_DSC_via_WebService()`/`DSC_WebService`. 
Here we only show how to deploy a clusterer, but a DSD can be published 
in a similar manner. Larger workflows can be created using `DST_Runner` from `stream`.

`streamConnect` uses the package `plumber` to manage web services.
The data is transmitted in serialized form. The default serialization format 
it `csv` (comma separated values). Other formats are `json` and `rds` 
(see `plumber::serializer_csv`). 


## Server: Create a Web Service
Creating a clustering web service process listening for data on the port.
```{r eval = TRUE}
library(streamConnect)

rp1 <- publish_DSC_via_WebService("DSC_DBSTREAM(r = .05)", port = port)
rp1
```

## Client: Connect to the Web Service
Connect to the web service with a local DSC interface.

```{r eval = TRUE}
library(streamConnect)

dsc <- DSC_WebService(paste0("http://localhost", ":", port), 
                      verbose = TRUE, config = httr::verbose(info = TRUE))
dsc
```
Note that the verbose output can help with debugging connection issues.


Cluster some data.
```{r eval = TRUE}
dsd <- DSD_Gaussians(k = 3, d = 2, noise = 0.05)

update(dsc, dsd, 500)
dsc


get_centers(dsc)
get_weights(dsc)

plot(dsc)
```

## Server: Stop the Web Service
Kill the web service process.
```{r eval = TRUE}
rp1$kill()
```

# Deploying the Server/Web Service

Web services and the socket-based server can be easily deployed to any server
or cloud system including containers. 
Make sure R and the package `streamConnect` and all dependencies are
installed. Create a short R script to start the server/service and deploy it.

```{r, eval = FALSE}
library(streamConnect)
port = 8001

publish_DSC_via_WebService("DSC_DBSTREAM(r = .05)", port = port, 
                           background = FALSE)
```

Web services can also be deployed using a plumber task file. The following call
does not create a server, but returns the name of the task file. 

```{r, eval = FALSE}
publish_DSC_via_WebService("DSC_DBSTREAM(r = .05)", serve = FALSE)
```

Open the file in R studio to deploy it or
read the [plumber Hosting vignette](https://www.rplumber.io/articles/hosting.html).

