---
title: "Getting Started with streamConnect"
author: "Michael Hahsler"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with streamConnect}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>")
```

`streamConnect` connects data stream sources and stream mining algorithms
through sockets or web services. It works with the `DSD` (data stream) and
`DSC` (data stream clustering) interfaces from the [stream package](https://github.com/mhahsler/stream).

This guide shows two common workflows: reading a stream over a socket and
updating a remote clusterer through a web service. For more control over the
web service endpoints and deployment, see the help pages for
`publish_DSD_via_WebService()` and `publish_DSC_via_WebService()`.

## Install and load

Install the released package from CRAN:

```{r install, eval=FALSE}
install.packages("streamConnect")
```

Load the package and select an available local port for the examples:

```{r load}
library(streamConnect)

port <- httpuv::randomPort()
port
```

## Read a data stream over a socket

Start a background process that publishes a stream. The publisher opens a
socket server and writes data in blocks. Since the socket format does not
include column names, provide them when creating the reader.

```{r socket-publish}
socket_process <- publish_DSD_via_Socket(
  stream::DSD_Gaussians(k = 3, d = 2),
  port = port
)
socket_process
```

Connect a local `DSD` reader and request points from the stream:

```{r socket-read}
socket_dsd <- DSD_ReadSocket(
  port = port,
  col.names = c("x", "y", ".class")
)

points <- stream::get_points(socket_dsd, n = 10)
head(points)
```

Close the reader and stop the publisher when finished:

```{r socket-stop}
stream::close_stream(socket_dsd)
if (socket_process$is_alive()) socket_process$kill()
```

## Update a remote clusterer over a web service

Web services also support request and response operations. Start a clusterer
in a background process, then create a local `DSC` interface for it:

```{r webservice-start}
web_port <- httpuv::randomPort()
clusterer_process <- publish_DSC_via_WebService(
  "DSC_DBSTREAM(r = .05)",
  port = web_port
)

clusterer <- DSC_WebService(paste0("http://localhost:", web_port))
clusterer
```

Send a batch of points to the remote clusterer and retrieve its centers and
weights through the local interface:

```{r webservice-update}
stream <- stream::DSD_Gaussians(k = 3, d = 2, noise = 0.05)
stats::update(clusterer, stream, n = 500)

stream::get_centers(clusterer)
stream::get_weights(clusterer)
```

Stop the service process when the work is complete:

```{r webservice-stop}
if (clusterer_process$is_alive()) clusterer_process$kill()
```

## Publish a stream through a web service

To expose a data source over HTTP, publish a `DSD` and read it with
`DSD_ReadWebService()`. The service returns CSV by default; JSON and RDS
serialization are also available through the `serializer` argument.

```{r webservice-stream, eval=FALSE}
port <- httpuv::randomPort()
source_process <- publish_DSD_via_WebService(
  "DSD_Gaussians(k = 3, d = 2)",
  port = port
)

remote_stream <- DSD_ReadWebService(paste0("http://localhost:", port))
stream::get_points(remote_stream, n = 10)

source_process$kill()
```

## Next steps

- Use `publish_DSD_via_Socket()` and `DSD_ReadSocket()` for continuous
  point-to-point stream transfer.
- Use `publish_DSD_via_WebService()` and `DSD_ReadWebService()` to serve
  stream data over HTTP.
- Use `publish_DSC_via_WebService()` and `DSC_WebService()` to update and
  inspect a remote clusterer.
- Set `serve = FALSE` to write a plumber task file for deployment with a
  hosting environment of your choice.

See the [package reference](https://mhahsler.r-universe.dev/streamConnect)
for the full function documentation.
